---
title: "SS3: User Guide: ICES fishing opportunites"

author: 
 - Henning Winker (GFCM - ICES WKBSS3) 

output:
  pdf_document:
    number_sections: yes
    toc: yes
    keep_tex: false
vignette: >
  %\VignetteIndexEntry{FLRef Vignetted}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r wrap-hook, echo = FALSE,message=F,warning=F}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```


```{r, echo = FALSE,message=F,warning=F}
knitr::opts_chunk$set(collapse = TRUE, comment = "  " ,fig.align = 'center', cache=FALSE,tidy.opts=list(width.cutoff=60), tidy=TRUE)

```




Load `r4ss`, `FLRef` and `ss3om`

```{r,warning=F,message=F}

library(r4ss)
library(FLRef)
library(ss3om)
library(ss3diags)

```

# SS3 assessment summary for anf.27.3a4


## Step 1: Set up file paths and folders structure for loading and saving the SS3 model

Set up the file path to the folder where the SS3 model folder with run is located.


Define name of reference model folder with the SS3 model outputs

```{r}
model = "ref"
```



Create `.rds` stock  file name 

```{r}

stock.file = paste0("anf.27.3a4_",model,".rds")
```

Load reference model 

```{r,eval=F}

ss3rep = SS_output(file.path(model))

```

Create an `rdata` folder in the assessment model subdirectory.

```{r}
rdata = file.path("rdata")
dir.create(rdata,showWarnings = FALSE)

```

Save the model as `rdata` file

```{r,eval=F}
saveRDS(ss3rep,file= file.path("rdata",stock.file))

```

...or load directly as `.rdata` if these had been saved already

```{r,echo=T}
ss3rep = readRDS(file.path(rdata,stock.file))
```

## Specify benchmarks

```{r}
Fmsy =	0.137
Fpa	 = 0.215
Fupper =	0.174
Flower =	0.105
Fp0.5	= 0.215
Btrigger =	38604
Blim	= 25686
Bpa	= 35692
TAC = 11293  # 2023
  
```

Make `data.frame` that is compatible with `FLPar` in `FLR` 

```{r}
benchmarks = data.frame(params=c("Fmsy","Fpa","Fupper","Flower","Fp0.5","Btrigger","Blim","Bpa"), data=c(Fmsy,Fpa,Fupper,Flower,Fp0.5,Btrigger,Blim,Bpa))
```

Convert to `FLPar` 

```{r}
bms = as(benchmarks,"FLPar")
bms
```

## Step 2: Convert SS3 to `FLStockR`

First, the `ssmvln()` from `FLRef` is used to generate the stock trajectories with uncertainty using a Monte-Carlo to generate a large number of iterations from multivariate log-normal approximation of the variance-covariance estimates.


For the conventional ICES advice, it is important to extend the assessment horizon to the reference year+1 to plot $SSB$ and recruitment one-step-ahead ($y+1$) of $F$ and $Catch$.

```{r}
years = ss3rep$startyr:(ss3rep$endyr+1)

mvn = FLRef::ssmvln(ss3rep, Fref="Btgt",verbose=F,years=years)
```

 The option `Fref=Btgt`, and not `Fref=MSY` is chosen because the reference points were based on $B_{tgt} = B_{40}$, with a corresponding $F_{tgt}=F_{40}$.

This can be checked by

```{r}

mvn$Btgtref
```

Next the `mvn` object can easily converted into the `FLStockR` object

```{r}
stk = ss2FLStockR(mvn)
```


By default, the reference points for $F_{tgt}$ and $B_{tgt}$ are extracted together with `MSY`, `B0` and `R0`.

```{r}
stk@refpts

```

However, for the final advice plot only the agreed benchmarks should be shown.
This can be done by specifying `stk@refpts` as the `FLPar` object `bms` 

Here, the reference points $F_MSY$, $F_{pa}=F_{p0.5}$, $MSY B_{trigger}$ ($B_{trigger}$),$B_{pa}$ and $B_{lim}$ are selected for plotting    

```{r}
stk@refpts = bms[c("Fmsy","Fpa","Btrigger","Bpa","Blim")]
```

The option `osa=TRUE` allows to plot the one-step-ahead forecast for $SSB$ and recruitment through 2023, while omitting 2023 for $F$ and Catch.


```{r, warning=F,message=F, fig.height=7,fig.width=9, fig.cap = paste0("Comparison of estimated stock status trajectories.")}
plotAdvice(stk,osa=T)
```

Next make a `FLStockR` with iterations generated MVLN Monte-Carlo (default nsim = 1000) to depict uncertainty

```{r}
# with uncertainty
stki = ss2FLStockR(mvn,output="iters",thin=1) 
# assign benchmark reference points
stki@refpts = stk@refpts
```

Check

```{r, warning=F,message=F, fig.height=7,fig.width=9, fig.cap = paste0("Uncertainty of estimated stock status trajectories with associated reference points, with solid line depicting the median")}
plotAdvice(stki,osa=TRUE)
```

## Step 3: Make Advice plot of stock status indicators with uncertainty

The final advice plot seeks to provide a standard format for presenting stock status indicators that shows the exact maximum likelihood estimates from the model (`stk`) and depicts the uncertainty around those from the Monte-Carlo approach (stki).

The plotting code allows to specify the years shown along the x-axis by adjusting the option  `break=c(seq(years[1],tail(years,1),5),tail(years,1))` depending on the length of the time series (here every 5 years and the last year)

```{r, warning=F,message=F, fig.height=7,fig.width=9, fig.cap = paste0("Estimated stock status trajectories with associated reference points for the 2024 benchmark update of European hake in GSAs 17-18")}
# Name plot padv
padv = plotAdvice(FLStocks(CIs=stki,Fit=stk),osa=TRUE)+
   scale_fill_manual(values = c("salmon","black"))+
  scale_color_manual(values = c(0,"black"))+
  theme(legend.position = "none")+
  scale_x_continuous(breaks=c(seq(years[1]+1,tail(years,1),5),tail(years,1)))+ 
  theme(axis.text.x = element_text(size=8, angle=90,vjust=0.5))
padv
```

### Save `FLStockR` objects in `.rds` format to `rdata`

It is adviced to specify additional information in the `FLStockR` object before saving it.

Label the `FLStockR` object properly


```{r}
stk2 = stk
stk2@name = "anf.27.3a4"
stk2@desc = "2023, SS3, WKBSS3"
```


Note that `stk@name` will be used through this script to label file names! 



```{r}
saveRDS(stk2,file=file.path(rdata,paste0(stk2@name,"_stk2.rds")))
saveRDS(stki,file=file.path(rdata,paste0(stk2@name,"_stki.rds")))
 
```


## Step 5: Generate output trajectories for the ICES advice template   

First a folder `òutput`  is created to save the outputs (Tables, Figures).

```{r}
output.dir = file.path("output")
dir.create(output.dir,showWarnings = FALSE)
```



The output from `ssmvln` can now be directly converted in the ICES time series compatible format from SS3 model by

```{r}
icests = ss2ices(mvn)

```

```{r}
knitr::kable(icests,"pipe",
      align ="lccccc",
      caption=" Assessment summary. High and Low refer to 95% confidence intervals.")

```



The `timeseries`  can be saved as `.csv` files 

```{r}
write.csv(icests,file=file.path(output.dir,paste0(stk2@name,"_ts.csv")),row.names = F)
```


\newpage

# F-based forecasts with SS3 and FLR


This section introduces F-based forecasting with SS3 for multi-fleet models (also works for single fleets and seasonal models), which is based on the so-called apic F values ($F_{apic}$), whereas the choice of the reference $F$-basis and the associated reference points may differ from the $F_{apic}$ scale. For instance, GFCM and ICES, $F_{bar}$ (option 5) is the default option. It is therefore necessary to rescale $F$-basis to $F_{apic}$ for generating forecasts that are consistent with, e.g., $F_{tgt}$ or $F_{cur}$.     


$F_{apic}$ is used for good reason in forecasts in order to account for multi- fleet selectivity. Comparing the partial impacts selectivity pattern requires setting the instantaneous rate of fishing mortaly $F$ at comparable constant levels. For this purpose, it is important to consider that the definition of selectivity differs across regions (e.g. $Fbar$ or exploitation rate). With regards to temporal compatibility of partial fleet selectivity effects, $F_{bar}$ has the undesirable property that its scale depends on the pre-specified age range across which $F_a$ is averaged. For example, if $F_{bar}$ is set to ages 2-4 to represent the dominant age classes under the current selectivity regime, but the goal is to evaluate the effect of selecting fish only at age-5, a common $F_{bar}$ would result in disproportionately high $F_a$ on ages 5+. This is because $F_{bar}$ is computed for age ranges that are hardly selected for the definition $S_a$ = $F_a$/$max(F_a)$. For this reason, it is more straight forward to use $F_{apical}$ as the standardized quantity $F$ quantify to account for partial impacts of fleet selectivity. 

In the following, step-by-step guidelines are provided to setup an $F_{apic}$, so that it correctly corresponds to the $F_{bar}$ baseline for $F_{tgt}$ across multiple fleets and seasons.

## Step 1: Basic setup

In this a case, a folder with the reference model run is created and the model outputs are loaded with `r4ss::SS_output`

next set up the folder where the SS3 model folder with run is located


Load the assessment model


```{r}
ss3rep=readRDS(file.path("rdata",stock.file))
```



To organise the forecast outputs, first create a subfolder `forecast`

```{r}
forecast.dir = file.path(model,"forecast")

dir.create(forecast.dir,showWarnings = F)
```


A new helper function `SSnewrun` was added to `ss3diags` to easily create subfolders for the forecast scenarios. First a `Fmsy` reference folder is created


To this specify a new subfolder path, where to run the forecast for a "fixed" $F_{MSY}$ scenarios

```{r}
fmsydir = file.path(forecast.dir,"Fmsy")

```

Create new F forecast model folder. Note that the data and control file and ss.exe names need to be specified if these diverge from the defaults `data.ss`, `control.ss` and `ss3.exe`

```{r}
dat = "ang_dat.ss"
ctl = "ang_ctl.ss"
ss.exe = "ss3.exe"

SSnewrun(model=file.path(model),dat=dat,ctl=ctl,newdir=fmsydir,ss.exe="ss3.exe")
```

Now the forecast file can be read be read with `r4ss`

```{r}
fc <- SS_readforecast(file.path(fmsydir, "forecast.ss"),verbose = F)
```

## Step 2: Initial F exploitation calculations for Fapic forecast 

Extract the `$exploitation` output from the report file

```{r}
Fexp = ss3rep$exploitation
```

Importantly, the `annual_F` are scaled to the F-basis (here $F_{bar}$), whereas fleet specific $F$ values are always given as $F_{apic}$

Next compute the combined $F_{apic}$ generically across fleets

```{r}
Fexp$Fapic = apply(as.matrix(ss3rep$exploitation[,-c(1:6)]),1,sum,na.rm=T)
```

and aggregate across seasons, by taking the `mean` and not the `sum`.

```{r}
Fapic = aggregate(Fapic~Yr,Fexp,mean)
```

Next compute the corresponding annual $F_{bar}$ values from the `annual_F` 

```{r}
Fbar = aggregate(annual_F~Yr,Fexp,mean)
```

To work out exact ratio between $F_{apic}$ and $F_{bar}$ so that it is consistent with the benchmark calculations with ss3, it is necessary to extract the reference years for selectivity from the `forecast.ss` file. 

The information required for the average selectivity conditions can be found in the `forecast.ss` file under `$Bmark_years`. The third and fourth position define the time horizon for the average selectivity across fleet, a value of -999 (here) indicates that the whole time series is use, but more commonly averages are taken, e.g. over the last 3 years, which can be specified as -2 0 or 2020 2022. The following code attempts to compute this generically.


```{r}
endyr = ss3rep$endyr
if(fc$Bmark_years[3]< -90){
  nfc = length(min(ss3rep$exploitation$Yr+1):endyr) # excluded init year
} else { # if specified (e.g. -2, 0)
  nfc = fc$Bmark_years[4]-fc$Bmark_years[3]+1 
}
# Benchmark reference years
bmyrs = (endyr-nfc+1):endyr
bmyrs
```

NOTE: Other than here, it recommended to set the Bmark_years in the `forecast.ss` so that all quantities represent the mean last 3 years (i.e. -2).
The advantage is that this allows using be consistent with the default FLR settings. 

```{r}
Fratio = mean(Fapic$Fapic[Fapic$Yr%in%max(bmyrs)]/Fbar$annual_F[Fbar$Yr%in%max(bmyrs)])
Fratio
```
`Fratio` defines the ratio of $F_{apic}$ to $F_{bar}$ for the reference period 


Set $F_{MSY}$ to benchmark

```{r}
Fmsy = bms["Fmsy"][[1]]
```

This value is given as $F_{bar}$ and therefore needs to be transformed to $F_{apic}$

```{r}
Fmsy.apic = Fmsy*Fratio
Fmsy  # Fbar
Fmsy.apic
```
## Step 3:  Setting up the manual F forecast input structure

First, do some basic house keeping for the model structure. This is designed to work generically for any multi-fleet or seasonal structure

```{r}
nseas = length(unique(ss3rep$exploitation$Seas)) # number of seasons
fleets = unique(ss3rep$fatage$Fleet) # fleets
nfleets = length(fleets) # number of fleet
```

Next, the mean Fapic by fleet and season is calculated 

```{r}
# subset to benchmark years for selectivity
fexp = ss3rep$exploitation[ss3rep$exploitation$Yr%in%bmyrs,] 
fexp = cbind(fexp[,1:2],fexp[,-c(1:5)])[,-3] #><>  single fleet trick
# flip
fexp = reshape2::melt(fexp, id.vars = c("Yr", "Seas"),
                      variable.name = "Fleet", 
                      value.name = "Fapic")
tail(fexp)
```
The forecast file requires Fleet IDs not names. In the next step these are extracted and fleet names are converted in to Fleet IDs 

```{r}
fleet = data.frame(Fleet =ss3rep$FleetNames,ID=ss3rep$fleet_ID)
fexp$Fleet= fleet[match(fexp$Fleet,fleet$Fleet),2]
```

\pagebreak

Then, the relative proportions of $F_{apic}$ by fleet and season can be computed

```{r}
Fap = aggregate(Fapic~Seas+Fleet,fexp,mean)
Fap$prop = Fap$Fapic/sum(Fap$Fapic)*nseas
Fap
```

In the next step, status quo $F_{sq}$ for forecasting over the intermediate year(s) is defined. This can be relatively easily changed to intermediate catch years.  Here, the $F_{sq}$ is taken as $F_{2022}$ to account for the systematically decreasing trend, and the intermediate years are set to 1, account for 1 data lag year.


```{r}
# F status q
nfsq = 1
nint = 1
```

Compute the $F_{sq}$ as $F_{apic}$ vector by season and fleet 

```{r}

fsq = ss3rep$exploitation[ss3rep$exploitation$Yr%in%((endyr-nfsq+1):endyr),] 
fsq = cbind(fsq[,1:2],fsq[,-c(1:5)])[,-3]  #><>  single fleet trick
fsq = reshape2::melt(fsq, id.vars = c("Yr", "Seas"),
                     variable.name = "Fleet", 
                     value.name = "Fapic")
Fsq = aggregate(Fapic~Seas+Fleet,fsq,mean)
```

Now, the forecast horizon can be defined in the loaded `starter.ss` object `fc`.  Summary statistics on catch opportunities require a three year forecast horizon:

(1) Intermediate year based on $F_{sq}$ or $TAC$
(2) Implementation year with catch and $F$ outcomes
(3) One-step-ahead forecast of the $SSB$ response and recruitment, when spawning is set to 1st of January (default)

```{r}
fc$Nforecastyrs = 3
nfyrs = fc$Nforecastyrs
fyrs= endyr+c(1:nfyrs) 
```

The F-vector that is passed on the forecast file comprises the season/fleet structure replicates for `ninit` for $F_{sq}$ and the forecast years under  $F_{tgt}$ that is scaled to $F_{apic}$ by the `Fratio` and portioned by fleets.

```{r}
fvec = c(rep(Fsq$Fapic,nint),rep(Fmsy*Fratio*Fap$prop,nfyrs-nint))
```

\pagebreak

Given the fleet, season, intermediate year and forecast years structures, the forecast table for the `forecast.ss` file can finally be constructed.

```{r}
fc$ForeCatch = data.frame("Year"=rep(fyrs,each=nseas*nfleets),"Seas"=1:nseas,
                          "Fleet"=rep(fleets,each=nseas),
                          "catch_or_F"=fvec,
                          "Basis"=99)
tail(fc$ForeCatch,9)
```                          
Note that the `Basis` 99 specifies that $Fs$ are inputted, including $F_{sq}$ for the intermediate year.

However, it also possible to input the $TAC$ for the intermediate year. In cases of multi-fleet models it is advice to apportion the $TAC$ for each fleet based on the most recent catch proportions by fleet (e.g. last years). The catch proportions can be computed from the information in

```{r}
ss3rep$catch[ss3rep$catch$Yr%in%tail(years,4),]
```

In this case there is only one fleet so the $TAC$ can be assigned directly to 2024 by declaring the `Basis` as 2 for catch


First, the relative proportions of $Catches$ by fleet and season is computed

```{r}
Cexp = ss3rep$catch[ss3rep$catch$Yr%in%bmyrs,] 
```

```{r}
Cap = aggregate(Exp~Seas+Fleet,Cexp,mean)
Cap$prop = Cap$Exp/sum(Cap$Exp)*nseas
Cap
```

```{r}
fc$ForeCatch[fc$ForeCatch$Year==2023,"catch_or_F"] = TAC*Cap$prop 
fc$ForeCatch[fc$ForeCatch$Year==2023,"Basis"] = 2
fc$ForeCatch
```
Finally, the forecast options need to be adjusted for manual input

```{r}
fc$eof=TRUE 
fc$InputBasis = -1
```

and then the modified `starter.ss` file can be saved 

```{r,message=FALSE,warning=F}
SS_writeforecast(fc, file =file.path(fmsydir, "forecast.ss"),overwrite = T,verbose = F)
```

## Step 4: Running Fmsy forecasts with checks

In principle, the `Ftgt` can serve as a reference and the model does not have to be run if the goal is set up a number forecasts relative to $F_{tgt}$.

However, for illustration, the `Ftgt` forecast is run to check that the $F_{apic}$ will produce $F_{bar}$ estimates that are consistent with $F_{tgt}$.

To run

```{r,eval=F}
r4ss::run(fmsydir, skipfinished = F, show_in_console = T, exe=ss.exe)
```

After the run is finished (here under 2.5 min) the output can be loaded again.

```{r,eval=F}
fmsyrep = SS_output(fmsydir)
# safe as rdata
save(fmsyrep,file=file.path(rdata,"fwd_fmsy.rdata"))

```

```{r,echo=F}
load(file.path(rdata,"fwd_fmsy.rdata"))
```


For a quick check the `plotAdvice()` from `FLRef` can be used, but first the forecast needs to be converted into a "simplified" `FLStock` object, using the function `ssmvln`. 

Switch `addprj=TRUE` on to add the forecast horizon.

```{r,warning=FALSE}

mvn = FLRef::ssmvln(fmsyrep,Fref="Btgt",addprj = T,verbose=F)
stkf = ss2FLStockR(mvn)
```

Again, assign benchmarks

```{r}
stkf@refpts=stk@refpts

```

It can be readily seen that the $F_{apic}$ based $F_{MSY}$ forecast corresponds indeed to the $F_{MSY}$ estimate on $F_{bar}$ scale.  


```{r p5a, message=F,warning=F,fig.height=6.5,fig.width=8, fig.cap = paste0("Stock trajectories for basecase run and a $F_{tgt}$ forecast, relative to reference points")}

plotAdvice(window(stkf,start=1991),osa=T)+geom_vline(xintercept = 2022.5,linetype=2)
```



\pagebreak

## Step 5: Looping through fixed forecast scenarios

Set up the F-based forecasts for $F=0$, $F_{pa}=F_{p0.5}$, $FMSY_{low}$, $FMSY_{high}$

```{r}
Fs = c("Fsq","F0","Fpa","Flower","Fupper")
```

Add Fadv according ICES MSY approach of applying the advice rule.
First extract the $SSB$ for the intermediate year + 1. 

```{r}
fwdyr = max(years+1)
fwdyr
b = an(ssb(stkf)[,ac(fwdyr)])
b
```
Apply Advice rule

```{r}

Fadv = applyAR(b,btrigger = an(bms["Btrigger"]),fmsy=an(bms["Fmsy"]))
Fadv
```
Here, $F_{adv}=F_{MSY}$ given the $SSB$ is above the MSY $B_{trigger}$

Add Fadv scenrario to runs

```{r}
Fs = c(Fs,"Fadv")
```
Specify $F$ scenario values for forecasts

```{r}
frefs = rbind(bms,FLPar(F0=0.0001,Fadv=Fadv))[Fs[Fs!="Fsq"]]

frefs
```


Specify forecast folders

```{r}
fdirs = file.path(forecast.dir,paste0(Fs))
```

Loop through the process of modifying the `forecast.ss` file iteratively. 


```{r,eval=F}


for(i in 1:length(Fs)){
  # create model folder by copying from the fmsydir with forecast setup
  SSnewrun(model=fmsydir,dat=dat,ctl=ctl,newdir=fdirs[i],ss.exe = ss.exe)
  # Read Forecast file
  fc <- SS_readforecast(file.path(fdirs[i], "forecast.ss"))
  # Apply Ffrac
  # Create F forecast vector (generic)
  # F target
  if(Fs[i]!="Fsq"){
  ftgt = an(frefs[Fs[i]])  
  # scale to Fapic and apportion to fleets
  fvec = c(rep(ftgt*Fratio*Fap$prop,nfyrs-nint))
  } else {
  # Use Fsq 
  fvec = c(rep(Fsq$Fapic,nfyrs-nint))
  }
  # Creat F forecast table in forecast.ss
  fc$ForeCatch[fc$ForeCatch$year%in%fc$ForeCatch$year[-c(1:nint)],"catch_or_F"] = fvec
  SS_writeforecast(fc, file =file.path(fdirs[i], "forecast.ss"),overwrite = T)
  r4ss::run(fdirs[i], skipfinished = F, show_in_console = TRUE, exe=ss.exe)
}
```

Load all runs in one go with `SSgetoutput`, including the Fmsy run

```{r,eval=F}
Ffwd = SSgetoutput(dirvec = c(fdirs,fmsydir))
save(Ffwd,file=file.path(rdata,"fwdFs.rdata"))
```

Check that these can be loaded

```{r}
load(file=file.path(rdata,"fwdFs.rdata"),verbose=T)

```

Quick check and plot `FLRef`

```{r}

fstks = FLStocks(Map(function(x,y){
  out = FLRef::ssmvln(x,Fref="Btgt",verbose=F,run=y,addprj=T)
  out = ss2FLStockR(out)
  out@refpts =  stk@refpts # replace with benchmarks 
  return(out)
},x=Ffwd,y=as.list(c(paste0(Fs),"Fmsy"))))
names(fstks) = c(paste0(Fs),"Fmsy")

```

Make final pretty forecast plot for F scenarios

```{r , fig.height=6.5,fig.width=9, warning=F,message=F, fig.cap = paste0("Trajectories for the forecast scanarios under different fishing mortalities. The vertical dashed lines denote from left to right: reference year, implementation year for short-term forecasts and the catch advice year for the short-term forecasts ")}

pstks = FLStocks(c(FLStocks(Assessment=window(stk,end=ss3rep$endyr)),
                   window(fstks,start=ss3rep$endyr)))

# Name plot pffwd (pffwd for F-based and pcfwd for catch-based)
pffwd = plotAdvice(window(pstks,start=ss3rep$endyr-10),osa=T)+
  geom_vline(xintercept = c(ss3rep$endyr:(ss3rep$endyr+2)),linetype=2,linewidth=0.3)+
  scale_color_manual(values=c("black",ss3col(length(fstks))))+
 scale_x_continuous(breaks=seq(1900,2200,2))+
  theme(axis.text.x = element_text(size=8, angle=90,vjust=0.5))

pffwd
```

Add uncertainties


```{r}


fstksi = FLStocks(Map(function(x,y){
  out = FLRef::ssmvln(x,Fref="Btgt",verbose=F,run=y,addprj=T)
  out = ss2FLStockR(out,output="iters")
  out@refpts =  stk@refpts # replace with benchmarks 
  return(out)
},x=Ffwd,y=as.list(c(paste0(Fs),"Fmsy"))))
names(fstksi) = paste0(c(paste0(Fs),"Fmsy"),".CI")

```



```{r , fig.height=6.5,fig.width=9, warning=F,message=F, fig.cap = paste0("Trajectories of medians and 90\\% CIs for the forecast scanarios under different fishing mortalities. The vertical dashed lines denote from left to right: reference year, implementation year for short-term forecasts and the catch advice year for the short-term forecasts ")}



pstksi = FLStocks(c(FLStocks(
  Assessment=window(stki,end=ss3rep$endyr)),
  window(fstksi,start=ss3rep$endyr)))

pffwdi = plotAdvice(window(pstksi,start=ss3rep$endyr-10),osa=T)+
  geom_vline(xintercept = c(ss3rep$endyr:(ss3rep$endyr+2)),linetype=2,linewidth=0.3)+
  scale_color_manual(values=c("black",ss3col(length(fstks))))+
  scale_fill_manual(values=c("darkgrey",ss3col(length(fstks))))+
 scale_x_continuous(breaks=seq(1900,2200,2))+
  theme(axis.text.x = element_text(size=8, angle=90,vjust=0.5))

pffwdi
```



## Step 6: Search for limit MSY Btrigger and Blim

Here, an approach is introduced to use `ss3om` to translate SS3 to a full `FLR` stock object to then apply `FLasher` for solving for the $F's$ that lead to $MSY B_{trigger}$ and $B_{lim}$ 

To do this the SS3 model is loaded to `FLR` with `ss3om`

```{r,message=F,warning=F}
run = "ref"
ss3stk = window(ss3om::readFLSss3(run,wtatage = TRUE))
sr = ss3om::readFLSRss3(run)
stk@name = "anf.27.3a4"
stk@desc = "2022,  ICES, SS3"
```
Convert to `FLStockR` and assign refpts

```{r}
ss3stk = FLStockR(ss3stk)
ss3stk@refpts = stk@refpts
```

Compare with ss3

```{r, warning=F,message=F, fig.height=7,fig.width=9, fig.cap = paste0("Comparison of estimated stock status trajectories from SS3 and ss3om")}
plotAdvice(FLStocks(SS3=window(stk,end=ss3rep$endyr),ss3om=ss3stk))
```

Find $F$ for $MSY B_{trigger}$ in 2025 using the new `FLRef` function `fwdF4B()`

```{r}
fwdBtri = fwdF4B(ss3stk,sr=sr,btgt=bms["Btrigger"],
                 niy=1,nfy=3,ival=TAC,imet="TAC",verbose=F)

F.Btri = an(tail(fbar(fwdBtri),1))
F.Btri


```

Find $F$ for $B_{lim}$


```{r}
fwdBlim = fwdF4B(ss3stk,sr=sr,btgt=bms["Blim"],
                 niy=1,nfy=3,ival=TAC,imet="TAC",verbose=F)

F.Blim = an(tail(fbar(fwdBlim),1))
F.Blim


```
Check

```{r, warning=F,message=F, fig.height=7,fig.width=9, fig.cap = paste0("Forecast of F leading to Blim and Btrigger in 2025")}
plotAdvice(FLStocks("F(Blim)"=window(fwdBlim,start=2000),"F(Btri)"=window(fwdBtri,start=2000)))
```

Run the F.Btri and F.Blim forecasts in SS3


```{r}
Flims = c("F.Btri","F.Blim")
frefs = rbind(frefs,FLPar(F.Btri=F.Btri,F.Blim=F.Blim))
```

Specify forecast folders

```{r}
fldirs = file.path(forecast.dir,paste0(paste0(Flims)))
```

Loop through the process of modifying the `forecast.ss` file iteratively. 


```{r,eval=F}


for(i in 1:length(Flims)){
  # create model folder by copying from the fmsydir with forecast setup
  SSnewrun(model=fmsydir,dat=dat,ctl=ctl,newdir=fldirs[i],ss.exe = ss.exe)
  # Read Forecast file
  fc <- SS_readforecast(file.path(fldirs[i], "forecast.ss"))
  # Apply Ffrac
  # Create F forecast vector (generic)
  # F target from fref FLPar
  ftgt = an(frefs[Flims[i]]) 
  
  # scale to Fapic and apportion to fleets
  fvec = c(rep(ftgt*Fratio*Fap$prop,nfyrs-nint))
  
  # Creat F forecast table in forecast.ss
  fc$ForeCatch[fc$ForeCatch$year%in%fc$ForeCatch$year[-c(1:nint)],"catch_or_F"] = fvec
  SS_writeforecast(fc, file =file.path(fldirs[i], "forecast.ss"),overwrite = T)
  r4ss::run(fldirs[i], skipfinished = F, show_in_console = TRUE, exe=ss.exe)
}
```

## Step 7: Add fixed TAC forecast

```{r}
fldirs = file.path(forecast.dir,paste0(paste0(Flims)))
```

Loop through the process of modifying the `forecast.ss` file iteratively. 


```{r,eval=F}
tacdir = file.path(forecast.dir,"TACsq")

  # create model folder by copying from the fmsydir with forecast setup
  SSnewrun(model=fmsydir,dat=dat,ctl=ctl,newdir=tacdir,ss.exe = ss.exe)
  # Read Forecast file
  fc <- SS_readforecast(file.path(tacdir, "forecast.ss"))
  # Creat Catch forecast table in forecast.ss
  fc$ForeCatch[,"catch_or_F"] = rep(TAC*Cap$prop,nfyrs) 
  fc$ForeCatch[,"basis"] = rep(2,nrow(fc$ForeCatch))
  SS_writeforecast(fc, file =file.path(tacdir, "forecast.ss"),overwrite = T)
  r4ss::run(tacdir, skipfinished = F, show_in_console = TRUE, exe=ss.exe)

```


\pagebreak

## Step 8 : Summarize ICES Fishing Opportunities

Load all runs in one go with `SSgetoutput`, including the Fmsy run

```{r,eval=F}
Ffwd2 = SSgetoutput(dirvec = c(fdirs,fmsydir,fldirs,tacdir))
save(Ffwd2,file=file.path(rdata,"fwdFs2.rdata"))
```

Check that these can be loaded

```{r}
load(file=file.path(rdata,"fwdFs2.rdata"),verbose=T)

```
Convert to `FLR`

```{r}

fstks = FLStocks(Map(function(x,y){
  out = FLRef::ssmvln(x,Fref="Btgt",verbose=F,run=y,addprj=T)
  out = ss2FLStockR(out)
  out@refpts =  stk@refpts # replace with benchmarks 
  return(out)
},x=Ffwd2,y=as.list(c(Fs,"Fmsy",Flims,"TACsq"))))
names(fstks) = c(paste0(Fs),"Fmsy",Flims,"TACsq")

```

Rearrange 

```{r}
icesorder = c("Fadv","Flower","Fupper","F0","Fpa","Fmsy","F.Blim","F.Btri","Fsq","TACsq") 

fstks = fstks[icesorder]
```

Make final pretty forecast plot for F scenarios


```{r , fig.height=6.5,fig.width=9, warning=F,message=F, fig.cap = paste0("Trajectories for the forecast scanarios under different fishing mortalities. The vertical dashed lines denote from left to right: reference year, implementation year for short-term forecasts and the catch advice year for the short-term forecasts ")}

pstks = FLStocks(c(FLStocks(Assessment=window(stk,end=ss3rep$endyr)),
                   window(fstks,start=ss3rep$endyr)))

# Name plot pffwd (pffwd for F-based and pcfwd for catch-based)
pffwd = plotAdvice(window(pstks,start=ss3rep$endyr-10),osa=T)+
  geom_vline(xintercept = c(ss3rep$endyr:(ss3rep$endyr+2)),linetype=2,linewidth=0.3)+
  scale_color_manual(values=c("black",ss3col(length(fstks))))+
 scale_x_continuous(breaks=seq(1900,2200,2))+
  theme(axis.text.x = element_text(size=8, angle=90,vjust=0.5))

pffwd
```


Create FLStocks with uncertainties


```{r}


fstksi = FLStocks(Map(function(x,y){
  out = FLRef::ssmvln(x,Fref="Btgt",verbose=F,run=y,addprj=T)
  out = ss2FLStockR(out,output="iters")
  out@refpts =  stk@refpts # replace with benchmarks 
  return(out)
},x=Ffwd2,y=as.list(c(Fs,"Fmsy",Flims,"TACsq"))))
names(fstksi) = c(paste0(Fs),"Fmsy",Flims,"TACsq")
fstksi = fstksi[icesorder]
```

Make ICES Fishing Opportunity Table input


```{r}
fish.ops = fwd2ices(stock=fstks,uncertainty=fstksi)

```



```{r}
knitr::kable(fish.ops,"pipe",
      align ="lccccc",
      caption=" Summary of Fishing Opportunities.")

```


The output can also be saved as `.csv` files 

```{r}
write.csv(fish.ops,file=file.path(output.dir,paste0(stk2@name,"_FishOps.csv")),row.names = F)
```
