---
title: "Install packages for stock synthesis benchmark"

author: 
 - Henning Winker (GFCM - WGAs) 

output:
  pdf_document:
    number_sections: yes
    toc: yes
    keep_tex: false
vignette: >
  %\VignetteIndexEntry{FLRef Vignetted}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r wrap-hook, echo = FALSE,message=F,warning=F}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```


```{r, echo = FALSE,message=F,warning=F}
knitr::opts_chunk$set(collapse = TRUE, comment = "  " ,fig.align = 'center', cache=FALSE,tidy.opts=list(width.cutoff=60), tidy=TRUE)

```

# Installation guidelines

## Install `FLR` packages 

### Basic packages `FLCore` and `ggplotFL` 

All `FLR` based packages are now available on r-universe (similar R cran), which means that for Windows users the packages are readily pre-compiled.

It is recommended to first install and test the two basic `FLR` packages `FLCore` and `ggplot` before proceeding with "heavier machinery" 

Before installing the packages clear your R environment

```{r,eval=FALSE,message=F,warning=F}
rm(list = ls())
```

and re-start R.

```{r,eval = F}

install.packages(c('FLCore', 'ggplotFL'), repo='https://flr.r-universe.dev')

```

Load packages

```{r,warning=F,message=F}
library(FLCore)
library(ggplotFL)

```

## Install `FLBRP`, `FLFishery` `FLasher` 

+ `FLBRP` solving for reference points in `FLR`

+ `FLasher` forecasting (requires `FLFishery`)

both packages use `C++` in the background. 

Please ensure that `FLash` is not loaded before installing or using `FLasher`. There are many conflicts.
Perhaps best to remove `FLash`.

```{r,eval=F}
install.packages(c('FLBRP','FLFishery','FLasher'), repo='https://flr.r-universe.dev')
```

Load packages

```{r}
library(FLBRP)
library(FLasher)

```

## `FLSRTMB`, `mse` and `FLRef` 

+ `FLSRTMB` for stock-recruitment fitting in TMB

+ `mse` many additional utilities and prerequisite for `FLRef` 

+ `FLRef `for advanced reference point estimation and producing advice plots

First, this needs some additional packages, first and foremost `devtools` and `TMB`

Note that it is recommended to install `TMB` from `type =  source`, which will also be required for running `spict`. The installation of `TMB` required that `Rtools` is correctly installed (see above)!

```{r, eval=F}
install.packages("devtools")
install.packages("TMB", type = "source")
```

Furthermore, best to install the `ggplot2` packages and reshape2

```{r,eval=F}
install.packages("ggplot2")
install.packages("reshape2")

```

Load packages

```{r,message=F,warning=F}
library(TMB)
library(ggplot2)
library(reshape2)

```

Now install `FLSRTMB`

```{r,eval=F}
install.packages(c('FLSRTMB'), repo='https://flr.r-universe.dev')
```

Next install `mse`

```{r,eval=F}
install.packages(c('mse'), repo='https://flr.r-universe.dev')
```

Load packages

```{r,message=F,warning=F}
library(FLSRTMB)
library(mse)

```

Install `FLRef`

```{r,eval=F}
install.packages(c('FLRef'), repo='https://flr.r-universe.dev')

```

Load package

```{r,message=F,warning=F}
library(FLRef)

```

\newpage 

## Install packages for Stock Synthesis

First in install `r4ss`, which is designed to load and evaluate ss3 models in R.
It is recommended to install the latest version of `r4ss` directly from github.

```{r, eval=FALSE}
devtools::install_github("r4ss/r4ss")
```

In addition, it is suggest to install the `FLR` package `ss3om`, which is needed to produce the partial F plot by gear/fleet

```{r,eval=F}
install.packages(c('ss3om'), repo='https://flr.r-universe.dev')

```

### FishLife

Let's install:

[FishLife](https://github.com/James-Thorson-NOAA/FishLife), which helps us to retrieve life-history estimates. 

Consider to update the [rfishbase](https://github.com/ropensci/rfishbase) version. During the process of installing the package you may be asked for. Please update it as needed. 

```{r }

#remotes::install_github("ropensci/rfishbase")
library("rfishbase")

```

You might face issues installing this package. Please run the below code before installing FishLife
 
```{r }

options(download.file.method = "wininet") 
 
```

The final step is confirming the package is properly installed by calling it. 
You should find the below loading information in your console

############################################################################

Loading package FishLife, developed by James Thorson for the National Marine Fisheries Service
For details and citation guidance, please see http://github.com/James-Thorson-NOAA/FishLife/

############################################################################

You also can retrieve the _vignettes_ to access the predictive values by running FishLife.
To retrieve the PDF please click in the ZIP folder and follow the path: **FishLife-3.1.0/vignettes/Access.pdf/**

Please note that a number of papers describing models, databases, and applications for stock assessment may be found under several headlines [Description of package, Further reading, and Applications for stock assessment](https://github.com/James-Thorson-NOAA/FishLife). 


```{r }

#devtools::install_github("james-thorson/FishLife", dep=TRUE)
library(FishLife)

```

### SPMpriors

[SPMpriors](https://github.com/Henning-Winker/SPMpriors), which provide a toolbox for generating priors for stock assessments from FishLife. Accordingly, _SPMpriors_ should be installed after _FishLife_ was. 
Note that the 'force' mode is being used in order to make sure _SPMpriors_ is installed and updated later than _FishLife_ is.  

```{r }

#devtools::install_github("henning-winker/SPMpriors",force = TRUE)
library(SPMpriors)

```

