---
title: "MSE for Southern seabass example"

author: 
 - Max Cardinale (SLU), modified from code 
 - provided by Iago Mosqueira for WKREBUILD and WKNEWREF

date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    number_sections: yes
    toc: yes
    #keep_tex: false
vignette: >
  %\VignetteIndexEntry{benchmark Vignetted}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
  
```{r wrap-hook, echo = FALSE,warning=F,message=F}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "  " ,fig.align = 'center', cache=FALSE,tidy.opts=list(width.cutoff=60), tidy=TRUE)

```

\pagebreak


# Southern bass example


Southern bass model has been developed in 2024 and reviewed by a benchmark.  


## Load models created by Create FLR stock objects for MSE.R file and attributes (created by MSE_reference points_estimation_best case.R file


Load libraries

```{r, message=FALSE, warning=FALSE}

library(mse)
library(FLRef)
library(FLBRP)
library(progressr)
#handlers(global=TRUE)
#install.packages(c("FLRef"), repo = "https://flr.r-universe.dev")
library(doFuture)
library(r4ss)
library(ss3diags)
library(ss3om)
library(icesTAF)
library(parallel)
library(mseviz)
library(dplyr)
#devtools::install_github("mebrooks/stockrecruit/StockRecruitSET", build_opts = c("--no-resave-data", "--no-manual"))
```

Get the libraries specifications

```{r, message=TRUE, warning=FALSE}

sessionInfo()
```


Define folder with R data files, models and other additional parameters (one basecase and one mock model (same as basecase) used as example)


```{r}
setwd("~/Max/Commitees/ICES/WKBBASS/MSE")

plan(multisession, workers=9)

its <- 500
fy  <- 2082
iy  <- 2024

basecase <- mget(load("Reference_run.rda"))
R0reduced <- mget(load("Reference_run.rda"))
load("Southerseabass_attributes.rda")

```

## OMs conditioning, defining FLStocks, FLSRs and SS3 refpts

```{r}

stks <- FLStocks(REF=basecase$stk, RED=R0reduced$stk)
srrs <- FLSRs(REF=basecase$srr, RED=R0reduced$srr)
srps <- list(REF=basecase$rps, RED=R0reduced$rps)
brps <- list(REF=basecase$brp, RED=R0reduced$brp)

```

Define functions

```{r}

getabSR <- function(stk, srr) {
  ab(fmle(as.FLSR(stk, model='bevholtSV'),
    fixed=list(s=params(srr)$s, v=params(srr)$v,
    spr0=params(srr)$v/params(srr)$R0)))
}
.
getBRPs <- function(stk, srr) {
  # COERCE FLSR as bevholt(a,b)
  nsr <- getabSR(stk, srr)

  # FIT brps
  brp <- brp(FLBRP(stk, sr=nsr))

  # EXTRACT brefpts
  brps <- remap(refpts(brp), R0=c('virgin', 'rec'), MSY=c('msy', 'yield'))
   
  return(brps)
}

```


### This part is only to hack Blim up to line 160


Load the stock object in FLR


```{r, message=FALSE, warning=FALSE, results='hide'}

stk_single <- readFLSss3(dir="C:/Users/mascar/Documents/Max/Commitees/ICES/WKBBASS/Southern stock/Basecase",wtatage = TRUE)
```


Load the SS model and parameters


```{r, results='hide', message=FALSE, warning=FALSE}

out <- SS_output(dir="C:/Users/mascar/Documents/Max/Commitees/ICES/WKBBASS/Southern stock/Basecase",covar=T, printstats=FALSE)
R0 <- exp(out$parameters$Value[out$parameters$Label=="SR_LN(R0)"])
s <- out$parameters$Value[out$parameters$Label=="SR_BH_steep"]
sigmaR <- out$parameters$Value[out$parameters$Label=="SR_sigmaR"]
rho <- out$parameters$Value[out$parameters$Label=="SR_autocorr"]
B0 <- out$derived_quants$Value[out$derived_quants$Label=="SSB_unfished"]
SSBcv <- out$derived_quants$StdDev[out$derived_quants$Label=="SSB_2023"]/out$derived_quants$Value[out$derived_quants$Label=="SSB_2023"]
Fcv <- out$derived_quants$StdDev[out$derived_quants$Label=="F_2023"]/out$derived_quants$Value[out$derived_quants$Label=="F_2023"]
BMSYss <- out$derived_quants$Value[out$derived_quants$Label=="SSB_MSY"]
FMSYss <- out$derived_quants$Value[out$derived_quants$Label=="annF_MSY"]
MSYss <- out$derived_quants$Value[out$derived_quants$Label=="Dead_Catch_MSY"]
TBOss <-  out$derived_quants$Value[out$derived_quants$Label=="Totbio_unfished"]

```


Coerce FLSR as bevholt(a,b) from SS


```{r}
nsr <- ab(fmle(as.FLSR(stk, model='bevholtSV'),
               fixed=list(s=s, v=B0,
                          spr0=B0/R0)))
```


Fit brps as single stock to hack Blim for the oms list. Blim set as 32% of B0


```{r}
brp_single <- brp(FLBRP(stk, sr=nsr))
brps_single <- remap(refpts(brp_single), R0=c('virgin', 'rec'), MSY=c('msy', 'yield'))
brps_single$Blim <- B0*0.15
```


\pagebreak


# Get Brps and add Blim to refpts


```{r}
brps <- lapply(srps, rbind, Blim=brps_single$Blim)
```


## Coerce SRRs to ab bevholt


```{r}
nsrs <- Map(function(x,y) getabSR(x, y),
  x=stks, y=srrs)
```


## Create FLoms

```{r}
oms <- Map(function(x,y,z) FLom(stock=x, sr=y, refpts=z),
  x=stks, y=srrs, z=brps)
```


## Extend to the future

```{r}
oms <- lapply(oms, function(x) propagate(fwdWindow(x, end=fy), its))
```


## Add SR deviances, same across OMs, and add those to OMs

```{r}
devs <- rlnormar1(its, sdlog=sigmaR, rho=rho, years=seq(2023, fy))

oms <- lapply(oms, function(x) {
  deviances(x) <- devs
  return(x)
  })
```


## F and SSB deviances for shortcut and STF


```{r}
sdevs <- shortcut_devs(oms[[1]], Fcv=Fcv, Fphi=0.423, SSBcv=SSBcv)
```

\pagebreak


# MP setup


## Setup standard ICES advice rule 


```{r}
arule <- mpCtrl(list(

  # (est)imation method: shortcut.sa + SSB deviances
  est = mseCtrl(method=shortcut.sa,
    args=list(SSBdevs=sdevs$SSB)),

  # hcr: hockeystick (fbar ~ ssb | lim, trigger, target, min)
  hcr = mseCtrl(method=hockeystick.hcr,
    args=list(lim=0, trigger=trigger, target=target,
    min=0, metric="ssb", output="fbar")),

  # (i)mplementation (sys)tem: tac.is (C ~ F) + F deviances
  # rec as GM ignoring last 2 years
  isys = mseCtrl(method=tac.is,
    args=list(recyrs=-2, fmin=0, Fdevs=sdevs$F))
  ))
```


Plotting the HCR


```{r, fig1, fig.height=4,fig.width=6, fig.cap = paste0("Harvest control rule used in the MSE"), warning=FALSE}
setwd("C:/Users/mascar/Documents/Max/Commitees/ICES/WKBBASS/MSE")
       plot_hockeystick.hcr(arule$hcr,
                            labels=c(trigger="Btrigger", target="Ftarget")) +
         xlab("SSB (t)") + ylab(expression(bar(F))) +
         ggtitle("HCR")

mseargs <- list(iy=iy-1, fy=fy, data_lag=1, management_lag=1, frq=1)
```


Get candidate values for Btrigger & Ftarget creating combinations of based on fraction of B0 (FBx ~ Bx * c(0.40, 0.45, 0.05)) and adding FMSY ~ BMSY * c(0.6, 0.8, 1) (add more combinations when running the final MSE)


```{r}

frps <- Map(function(x, y) {
    lapply(seq(40, 45, by=5), function(i) {
      Fbrp(computeFbrp(x, sr=y, proxy="bx", x=i, blim=0.15))
    })
  }, x=stks, y=nsrs)

opts <- Map(function(x, y) {
  res <- list(
    target=rep(unlist(lapply(x, '[', 1)), each=3),
    trigger=unlist(lapply(seq(0.40, 0.45, by=0.05),
      function(i) (c(refpts(y)$SB0) * i) * c(0.60, 0.8, 1))))
  res$target <- c(res$target, rep(c(refpts(y)$FMSY), 3))
  res$trigger <- c(res$trigger, c(refpts(y)$SBMSY) * c(0.6, 0.8, 1))
  return(res)
  }, x=frps, y=oms)

nms <- c(paste0(rep(paste0("FB", seq(40, 45, by=5)), each=3),
  rep(c(".6", ".8", "1"), 2)), c("FMSY.6", "FMSY.8", "FMSY1"))
```


Run for all options on 'hcr' control element for all OMs and ICES advice. To be done on a cluster pc and saved for loading (plans is a very large file)


```{r, eval=F}
system.time(plans <- Map(function(x, y) {

  
  runs <- mps(x, ctrl=arule, args=mseargs, hcr=y, names=nms)
  
  runs$advice <- advice <- mp(x, ctrl=arule, args=mseargs)

  return(runs)
  
  }, x=oms, y=opts))
```


Save the MSE results


```{r, eval=FALSE}
save(oms, plans, file="plans.rda", compress="xz")
```


## Compute MSE performances


Define performances metrics


```{r}
metrics <- list(SB = ssb, F = fbar, C = landings, TC = catch, Rec = rec)

stats <- list(
  medianFmsy= list(~yearMedians(F/FMSY), name="F/Fmsy",
    desc="Median annual F/Fmsy"),
  medianBmsy = list(~yearMedians(SB/SBMSY), name="B/Bmsy",
    desc="Median annual B/Bmsy"),
  medianCmsy = list(~yearMedians(C/MSY), name="Catch/MSY",
    desc="Median Catch/MSY over years"),
  aavC = list(~yearMedians(iav(C)), name="AAV",
    desc="Median annual variation in catches"),
  riskBlim = list(~apply(iterMeans((SB/Blim) < 1),1,max), 
    name="P3(B<Blim)", desc="Probability that SSB < Blim"),
  risk10SB0 = list(~apply(iterMeans((SB/(SB0 * 0.10)) < 1), 1, mean),
    name="P(B<SB0.10)", desc="Probability that SSB < 10% SB0"),
  P80BMSY = list(~apply(iterMeans((SB/(SBMSY * 0.8)) > 1), 1, max),
    name="B>80Bmsy", desc="Probability that SSB > 80% x Bmsy"),
  medianSBMSY = list(~yearMedians(SB/SBMSY), name="SSB/SSB[MSY]",
    desc="Median annual SSB/SSBmsy"),
  medianFMSY = list(~yearMedians(F/FMSY), name="F/F[MSY]",
    desc="Median annual F/FMSY")
)
```


Compute performances, add HCR parameters Btrigger and Ftarget and define long time horizon for evaluation


```{r, eval=FALSE}
plans_perf <- lapply(plans, function(x) {
  
  res <- performance(x, statistics=stats,
                     years=list(all=2024:fy, long=2053:fy))

  hps <- rbindlist(lapply(x, function(i)
    args(control(i)$hcr)[c("trigger", "target")]), idcol='mp')
  
  performance(x) <- merge(res, hps, by='mp')

  return(x)
})

```


Create MSE performance table


```{r, eval=FALSE}
perf <- rbindlist(lapply(plans_perf, performance), idcol="om")
write.csv(perf,file="msePerf_data.csv")
```


Save performances objects


```{r, eval=FALSE}
save(oms, plans, plans_perf, perf, file="plans.rda", compress="xz")
```


Load performances objects derived from a run made on a 16 cores cluster computer on Linux


```{r}
load("~/Max/Commitees/ICES/WKBBASS/MSE/plans.rda")
```


# MSE Performance plots and tables


```{r fig2, fig.height=12,fig.width=10, fig.cap = paste0("Southern seabass stock. MSE performance plot for OM1")}

ncol = length(unique(perf$mp)) # n colors
perfom1 = perf[perf$om=="REF",]
pbp = plotBPs(perfom1[perfom1$year=="long",],
statistics=c("medianFmsy","medianBmsy","medianCmsy", "aavC", "riskBlim", "P80BMSY"), 
size=3, target = c(medianFmsy=1,medianBmsy=1, medianCmsy=1),
limit= c(riskBlim=0.05,P80BMSY=0.95),
yminmax = c(0.05, 0.95))+theme_bw()+
facet_wrap(~name,scales = "free_y",ncol=2)+
ggtitle(paste0("Performance: Long"))+
ylab("Performance statistics")+
scale_fill_manual(values=ss3col(ncol))+ # USE FLRef::ss3col
theme(axis.text.x=element_blank())+xlab("Candidates")
pbp

```


```{r fig3, fig.height=12,fig.width=10, fig.cap = paste0("Southern seabass stock. MSE performance plot for OM2")}

ncol = length(unique(perf$mp)) # n colors
perfom2 = perf[perf$om=="RED",]
pbp = plotBPs(perfom2[perfom2$year=="long",],
statistics=c("medianFmsy","medianBmsy","medianCmsy", "aavC", "riskBlim", "P80BMSY"), 
size=3, target = c(medianFmsy=1,medianBmsy=1, medianCmsy=1),
limit= c(riskBlim=0.05,P80BMSY=0.95),
yminmax = c(0.05, 0.95))+theme_bw()+
facet_wrap(~name,scales = "free_y",ncol=2)+
ggtitle(paste0("Performance: Long"))+
ylab("Performance statistics")+
scale_fill_manual(values=ss3col(ncol))+ # USE FLRef::ss3col
theme(axis.text.x=element_blank())+xlab("Candidates")
pbp

```


```{r fig4, fig.height=12,fig.width=10, fig.cap = paste0("Southern seabass stock. MSE performance plot, ensemble")}

ncol = length(unique(perf$mp)) # n colors
pbp = plotBPs(perf[perf$year=="long",],statistics=c("medianFmsy","medianBmsy","medianCmsy", "aavC", "riskBlim", "P80BMSY"), size=3,
              target = c(medianFmsy=1,medianBmsy=1, medianCmsy=1),
              limit= c(riskBlim=0.05,P80BMSY=0.95),
              yminmax = c(0.05, 0.95))+theme_bw()+
  facet_wrap(~name,scales = "free_y",ncol=2)+
  ggtitle(paste0("Performance: Long"))+
  ylab("Performance statistics")+
  scale_fill_manual(values=ss3col(ncol))+ # USE FLRef::ss3col
  theme(axis.text.x=element_blank())+xlab("Candidates")
pbp

```

\pagebreak


## MSE Table


```{r, warning=FALSE, message=FALSE}

perf_table <- perf %>%                              # Specify data frame
  group_by(om, name, mp, year, trigger, target) %>%     # Specify group indicator
  summarise_at(vars(data),                          # Specify column
               list(Median = median),na.rm = TRUE)  # Specify function

write.csv(perf_table,file="msePerf_table.csv")
```


## MSE kobe plot

```{r fig5, fig.height=8,fig.width=10, fig.cap = paste0("Southern seabass stock. MSE kobe plot, ensemble"), warning=FALSE}

kbcex =function(){theme(plot.title = element_text(size=10),
                        legend.key.size = unit(0.3, 'cm'), #change legend key size
                        legend.key.height = unit(0.4, 'cm'), #change legend key height
                        legend.key.width = unit(0.4, 'cm'), #change legend key width
                        legend.text = element_text(size=10)) #change legend text font size
}
kobeMPs(perf,y="medianFmsy", x="medianBmsy", SBlim=NULL, Ftarget = 1)+
  ylab(expression(F/F[MSY]))+xlab(expression(B/B[MSY]))+ylim(0,2.5)+kbcex()+theme()

```

\pagebreak


## Trajectories plot for each OM

```{r fig6, fig.height=10,fig.width=10, fig.cap = paste0("Southern seabass stock. Trajectories plot for OM1")}
plot(oms[[1]], plans[[1]])

```

\pagebreak


```{r fig7, fig.height=10,fig.width=10, fig.cap = paste0("Southern seabass stock. Trajectories plot for OM2")}
plot(oms[[2]], plans[[2]])

```

\pagebreak

## Plotting single trajectories and medians for each OMs

```{r}
rp = oms$REF@refpts
om1 = window(FLStockR(oms[[1]]@stock),end=2023)
om1@refpts = FLPar(Fmsy=rp["FMSY"],Bmsy=rp["SBMSY"],Blim=rp["Blim"],MSY=rp["MSY"])
stks1 = FLStocks(lapply(plans[[1]],function(x){
  out = FLStockR(x@om@stock)
  out@refpts = om1@refpts
  out
}))

res1 = FLStocks(c(FLStocks(om1),stks1))
names(res1)[1] = "om1"

# Medians
med1 = FLStocks(lapply(res1,function(x){stockMedians(x)}))
```

```{r}
rp = oms$RED@refpts
om2 = window(FLStockR(oms[[2]]@stock),end=2023)
om2@refpts = FLPar(Fmsy=rp["FMSY"],Bmsy=rp["SBMSY"],Blim=rp["Blim"],MSY=rp["MSY"])
stks2 = FLStocks(lapply(plans[[2]],function(x){
  out = FLStockR(x@om@stock)
  out@refpts = om1@refpts
  out
}))

res2 = FLStocks(c(FLStocks(om2),stks2))
names(res2)[1] = "om2"

# Medians
med2 = FLStocks(lapply(res2,function(x){stockMedians(x)}))
```


```{r fig8, fig.height=10,fig.width=10, fig.cap = paste0("Southern seabass stock. Trajectories plot for OM1 medians"), warning=FALSE}

plotAdvice(med1)+geom_line(linewidth=0.8)+
  scale_color_manual(values=c("black",sscol(length(stks1))))+
  scale_x_continuous(breaks=seq(1960,fy,5))+
  theme(axis.text.x = element_text(size=8, angle=90))+
  guides(col = guide_legend(ncol = 1))

```

\pagebreak

```{r fig9, fig.height=10,fig.width=10, fig.cap = paste0("Southern seabass stock. Trajectories plot for OM2 medians"), warning=FALSE}

plotAdvice(med2)+geom_line(linewidth=0.8)+
  scale_color_manual(values=c("black",sscol(length(stks2))))+
  scale_x_continuous(breaks=seq(1960,fy,5))+
  theme(axis.text.x = element_text(size=8, angle=90))+
  guides(col = guide_legend(ncol = 1))

```

\pagebreak

```{r fig10, fig.height=10,fig.width=10, fig.cap = paste0("Southern seabass stock. Trajectories plot for OM1, all"), warning=FALSE}

plotAdvice(res1)+
  scale_color_manual(values=c("black",sscol(length(stks1))))+
  scale_fill_manual(values=c("black",sscol(length(stks1))))+
  scale_x_continuous(breaks=seq(1960,fy,5))+
  theme(axis.text.x = element_text(size=8, angle=90))+
  guides(col = guide_legend(ncol = 1))

```


```{r fig11, fig.height=10,fig.width=10, fig.cap = paste0("Southern seabass stock. Trajectories plot for OM2, all"), , warning=FALSE}

plotAdvice(res2)+
  scale_color_manual(values=c("black",sscol(length(stks2))))+
  scale_fill_manual(values=c("black",sscol(length(stks2))))+
  scale_x_continuous(breaks=seq(1960,fy,5))+
  theme(axis.text.x = element_text(size=8, angle=90))+
  guides(col = guide_legend(ncol = 1))

```

\pagebreak


## Select best HCR target and trigger combination

```{r}
perfCsv <- read.csv("msePerf_table.csv")
#OM <- c("REF", "RED")
OM <- c("REF")

p05 <- subset(perfCsv, om %in% OM & (name == "P3(B<Blim)" & Median < 0.05))
catchP05 <- subset(perfCsv, om %in% OM & (name == "Catch/MSY" & mp %in% p05$mp))
bestCase <- subset(catchP05, Median == max(Median))
bestCase <- bestCase %>%
  select(mp, trigger, target)
bestCase$target <- round(bestCase$target, 3)
names(bestCase)[1] <- "MP"
names(bestCase)[2] <- "Btrigger"
names(bestCase)[3] <- "Ftarget"

knitr::kable(bestCase,"pipe",
      align ="lc",caption="Summary of estimated reference 
      points for reference case model of Southern seabass stock")
```


# Estimate Blim with different methods as the minimum SSB level that resulted in a recruitment higher that the median and fractions of B0 and different fractions of B0

```{r}
library(StockRecruitSET)

flsr <- as.FLSR(stk_single)
S <- an(ssb(flsr))
R <- an(rec(flsr))

Bpaemp <- calcBlim(S, R, quant = 0.5, type = 1, nmin=1, AIC=TRUE)
Blim_emp = round(Bpaemp/(exp(1.645*SSBcv)),0)
Blim_emp_defCV = round(Bpaemp/(exp(1.645*0.2)),0) 
Blim24 <- round(B0*0.24,0)
Blim30 <- round(B0*0.30,0)

###Bloss calculations
Bloss = min(ssb(stk_single))
BpaBloss= Bloss
BlimBloss = BpaBloss/(exp(1.645*SSBcv))


Bpaemp
Blim_emp
Blim_emp_defCV
Blim24
Blim30
Bloss
BpaBloss
BlimBloss

```
